---
title: "Women's Soccer Data"
output: html_notebook
---

We'll use this notebook to explore some NCAA women's soccer data and [our questions](https://docs.google.com/spreadsheets/d/1kTSrnl1URmcexIWLDMWqDnp3q-c2OS45uifnhjp10n4/edit#gid=0). First, let's load the tidyverse library:

```{r}
library(tidyverse)
```

We'll want to start with 2023 match data, which is located in [this GitHub repository](https://github.com/dwillis/NCAAWomensSoccerData). We can load the match data directly from GitHub using `read_csv`:


```{r}
matches_2023 <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensSoccerData/main/data/ncaa_womens_soccer_matchstats_2023.csv")

matches_2022 <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensSoccerData/main/data/ncaa_womens_soccer_matchstats_2022.csv")
```

Let's take a look at this data to see what it contains, using `head` and `glimpse`:

```{r}
head(matches_2023)
glimpse(matches_2023)
```

Now, onto the questions using aggregates, which means `group_by` and `summarize`:

1. How many shots in total has Maryland attempted and how many have been on goal? Calculate a percentage of shots on goal.

```{r}
matches_2023 |> 
  group_by(team) |> 
  summarise(
    total_shots = sum(sh_att), 
    total_on_goal = sum(so_g),
    percent = (total_on_goal/total_shots)*100
  )
```

2. What about the same shot stats from last season's first four games? You'll need to load the matches from 2022 - just change the season in the URL - and filter by date to find the first four games.

```{r}
matches_2022 |> 
  filter(team == 'Maryland Terrapins, Terps') |> 
  group_by(team) |> 
  summarise(
    total_shots = sum(sh_att), 
    total_on_goal = sum(so_g),
    percent = (total_on_goal/total_shots)*100
  ) |> 
  arrange(desc(percent))
```


3. What about draws - Maryland started 2022 with 4 ties. Have any other teams had as many this season?

```{r}
matches_2023 |> 
  filter(outcome == 'Draw') |> 
  group_by(team) |> 
  summarise(
    draws = n()
  ) |> 
  arrange(desc(draws)) |> 
  filter(draws > 3)
```

In addition to Maryland, only eight other teams this season have drawn four games. The Terps are the only Power 5 school to have done so.

4. How many total corner kicks has Maryland had so far and how many have its opponents have? Calculate the difference between the two totals for every team.

```{r}

matches_2023 |> 
  group_by(team) |> 
  summarise(
    total_corners = sum(corners), 
    total_opp_corners = sum(defensive_corners),
    diff = (total_corners-total_opp_corners)
  ) |> 
  arrange(desc(diff))
```
Maryland is -2 in corner differential this season. Troy is absolutely crushing it on corners.


### Significance tests

Make sure you load the 2022 and 2023 matches data above. For both, we'll need to add a column called `total_goals` and populate it using `mutate()`. You'll save that new dataframe with the same name:

```{r}
matches_2022 <- matches_2022 |> mutate(total_goals = goals + defensive_goals)
matches_2022_partial <- matches_2022 |> filter(date < '2022-09-13')
matches_2023 <- matches_2023 |> mutate(total_goals = goals + defensive_goals)

matches_2022_partial |> summarize(sum(total_goals))

matches_2023 |> summarize(sum(total_goals))

```

Our Null Hypothesis is that there is no significant difference in the average number of goals scored in 2022 and 2023. Our Alternative Hypothesis is there is a significant difference in the average number of goals scored in 2022 and 2023.

```{r}

# Perform a t-test to compare average goals in 2022 and 2023
result <- t.test(matches_2022$total_goals, matches_2023$total_goals)

# Print the t-test result
print(result)

```
Based on the output (read from the bottom), is scoring up this year, and is that result significant? Explain your answers.



Second question: is this a fair comparison? If not, what would make it fairer?


```{r}
matches_2022_early <- matches_2022 |> filter(date < '2022-09-14')
t.test(matches_2022_early$total_goals, matches_2023$total_goals)
```


### Defensive model

```{r}
matches_2023 <- matches_2023 |> 
  mutate(differential = team_score - opponent_score)
```

```{r}
fit2 <- lm(differential ~ so_g + defensive_so_g , data = matches_2023)
summary(fit2)
```
```{r}
matches_2023 <- matches_2023 |> mutate(predicted = predict(fit2), residuals = residuals(fit2))


big_ten = c('Wisconsin Badgers', 'Maryland Terrapins, Terps', 'Michigan Wolverines', 'Rutgers Scarlet Knights', 'Indiana Hoosiers', 'Iowa Hawkeyes', 'Penn St. Nittany Lions', 'Minnesota Golden Gophers', 'Illinois Fighting Illini', 'Ohio St. Buckeyes', 'Northwestern Wildcats', 'Michigan St. Spartans', 'Purdue Boilermakers')

matches_2023 |> filter(differential < 5, team %in% big_ten, outcome == 'Draw') |> arrange(residuals) |> select(date, team, opponent, home_away, outcome, team_score, opponent_score, so_g, defensive_so_g, predicted, residuals)
```

